
```{r}
#libraries
install.packages('googleVis')
install.packages('ggplot2')
install.packages('dplyr')
install.packages('ggplot2')
library(dplyr)
library(tidyverse)
library(ggplot2)
```
A1. Investigating the Distribution of Indigenous Australians



1. Use R to read, wrangle and analyse the data in Data1. Make sure you describe any
complications you encounter and the steps you take when answering the following questions
```{r }
#loading the data1
data1<-read.csv("FIT5145_S2_2020_A1_Data1_IndigAusPopData_byregion.csv",fileEncoding="UTF-8-BOM")
head(data1)

```
The above data displays the Indigenous population By Region
Here on examing the data I can clearly say that the Age column has around 1000 wrong values due to some human error where the age group should be [10-14] it is mistakenly written as Oct-14[1-10-14], same goes with 5-Sep i.e.age group [5-9].
The name of the 1st and 2nd columns both are same which creates redundancy and confusion so changing INDIGENOUS_REGION to Region_Code will make it easier for user to read and understand data.


```{r}
#changing the name of the first column using rename method
data_clean1<-data1 %>% 
  rename(Region_Code=INDIGENOUS_REGION,Year=TIME)
  
#converting age from factor to character type 
  data_clean1$Age <- as.character(data_clean1$Age)
  

# transforming the data of Age column
  data_clean1$Age[data_clean1$Age == "Oct-14" ] <- "10 - 14"
  data_clean1$Age[data_clean1$Age == "5-Sep" ] <- "5 - 9"
  
#converting age to factor again  
  data_clean1$Age <- as.factor(data_clean1$Age)
  
head(data_clean1)
main<-data_clean1

```

a. What regions have the maximum and minimum total Indigenous populations in 2016 and
2031?

```{r}

Indigenous_Pop<-data_clean1%>%
    group_by(Indigenous.region,Year) %>%
    filter(Year =='2016' || Year =='2031')%>%
    summarise(Total_Indigenous_Population = sum(Value, na.rm = TRUE)) %>%
    arrange(desc(Total_Indigenous_Population))
Indigenous_Pop

```
Dataframe:-
In this case, The dataframe Indigenous_Popis created which uses group_by function on Indigenous.Region and TIME to extract only Region name and Time. Furthermore,  filter was applied to extract values only for the year 2016 and 2031 which was followerd by summarise function to sum all the values on the basis of Indigenous.region.

Task:-
The next task is to find the regions maximum and minimum total Indigenous populations in 2016 and 2031, For this scenario Year filter can be applied on the Indigenous_Pop dataframe to get the maximum and minimum values.


```{r}
value_2016<-Indigenous_Pop%>%
  filter(Year== '2016')

max_Indi_region<-value_2016[which.max(value_2016$Total_Indigenous_Population),]
max_Indi_region

  
```
The maximum total Indigenous populations in 2016 is: NSW Central and North Coast 85169

```{r}
min_Indi_region<-value_2016[which.min(value_2016$Total_Indigenous_Population),]
min_Indi_region
```
The minimum total Indigenous populations in 2016 is: Port Lincoln - Ceduna 2738

```{r}
value_2031<-Indigenous_Pop%>%
  filter(Year== '2031')

max_Indi_region_2031<-value_2031[which.max(value_2031$Total_Indigenous_Population),]
max_Indi_region_2031

```
The maximum Total Indigenous population for 2031 is Brisbane 129835
```{r}
min_Indi_region_2031<-value_2031[which.min(value_2031$Total_Indigenous_Population),]
min_Indi_region_2031
```
The minimum total Indigenous populations in 2031 is: Port Lincoln - Ceduna 2881

To recapitulate, the problem was straight forward ad it was not complicated at all, here group_by function on Region, Time  and filtering the Year column played an important role in extracting the maximum and minimum Total Indigenous population.



-----------------------------------------------------------------------------------------------------------------------------
b. What region/s have the maximum and minimum growth or decay rates of their total Indigenous population between 2016 and 2031?
-------------------------------------------------------------------------------------------------------------------------------


```{r}
growth_decay<-data_clean1%>%
    group_by(Indigenous.region,Year) %>%
    filter(Year =='2016' || Year =='2031')%>%
    summarise(Total_Population=sum(Value, na.rm = TRUE)) %>%
    mutate(rates = abs(Total_Population[2] - Total_Population[1])/(Total_Population[1])*100) %>%
    arrange(desc(rates))
growth_decay 



```
The growth_decay dataframe consist of Indigenous.region, Year ehich was derived using group_by function, The year column was filtered to 2016 and 2031 only.Moreover summarise function was used to calculate the total population by sumimg the value grouping by region and time.The process till here was similar to what was done in previous questions. 

However, calculating the rates was a bit tricky as I was not aware of mutate function earlier, but the point when I understood how it perform changes to all the rows with efficiency, I was comfortable using it. So in this part rates are calculated by subtracting the summarised value of succeeding row with the current  and dividing it by current row * 100. 

I even tried lead() and lag() functions but the results were not correct.

To find the maximum and minimum growth or decay rates of their total Indigenous population between 2016 and 2031, same approach is used as above.

```{r}
rate_2016<-growth_decay%>%
  filter(Year== '2016')

max_rate_2016<-rate_2016[which.max(rate_2016$rates),]
max_rate_2016

```
The region having maximum growth or decay rates of their total Indigenous population in 2016 is ACT  by 54.90483

```{r}

min_rate_2016<-rate_2016[which.min(rate_2016$rates),]
min_rate_2016
```

The region having minimum growth or decay rates of their total Indigenous population in 2016 is Port Lincoln - Ceduna  by 5.22279
```{r}
rate_2031<-growth_decay%>%
  filter(Year== '2031')

max_rate_2031<-rate_2031[which.max(rate_2031$rates),]
max_rate_2031

```
Similar to 2016, The region having maximum growth or decay rates of their total Indigenous population in 2031 is ACT  by 54.90483

```{r}
min_rate_2031<-rate_2031[which.min(rate_2031$rates),]
min_rate_2031
```

The region having minimum growth or decay rates of their total Indigenous population in 2031 is Port Lincoln - Ceduna  by 5.22279

---------------------------------------------------------------------------------------------------------------------------------
c

Plot and describe the growth or decay of the total Indigenous populations for the capitals of
the 8 state/territories across all time periods.

For these calculations, you will need to work out the growth/decay rates for each time period,
where the total population of the capital in time period N is compared to that in time period
N+1.

e.g., if 2017 population = 5050 and 2016 population = 5000,
then rate = (5050 - 5000) / 5000 = 50/5000 = 0.01, so 1% growth for 2016-2017
---------------------------------------------------------------------------------------------------------------------------------
To solve this problem, filter has been applied to extract all the capitals of 8 states like Perth for WA, Melbourne for Victora,
Darwin for NT, Sydney & Wollongong for NSW, Brisbane for QLD, Adelaide for SA. However the values for states like Canberra NT and Hobart for TAS were missing, so in that case the values present in the dataset i.e Tasmania and ACT was taken as capitals.The filter is used to perform this task and the data is stored in capitals dataframe



```{r}
capitals<-subset(data_clean1, Indigenous.region == "Brisbane" |Indigenous.region == "ACT" |Indigenous.region == "Perth"|Indigenous.region == "Melbourne"|Indigenous.region =="Sydney - Wollongong"|Indigenous.region == "Darwin"|Indigenous.region == "Adelaide"|Indigenous.region == "Tasmania")
capitals


```
After filtering the capitals,the growth and decay rates are calculated using lead() function.
Lead function points the value of its preceding column. Therefore here lead function takes the value of its preceding year and the calculates the differnce.The result is divided by the preeceding value and muliplied by 100 to get the rate. 

For example
7730 (2017)- 7513(2016)/7513(2016) *100 = 2.888

The results are applied to all the rows using mutate function.

```{r}
#dataframe
cap_growth_decay<-capitals%>%
  
  #group_by region name and time 
    group_by(Indigenous.region,Year) %>%
  
 # summarise  function to calculate Total Population
  summarise(Total_Population = sum(Value, na.rm = TRUE)) %>%

  #Calculate rates
  mutate(growth = (lead(Total_Population)-Total_Population)/Total_Population *100)

#time period
cap_growth_decay$year_group<-cut(cap_growth_decay$Year, breaks=c(2016, 2017, 2018,2019,2020,2021,2022,2023,2024,2025,2026,2027,2028,2029,2030,2031), right = FALSE)
cap_growth_decay
```
Calculating the rates was easy but to create the time period based on it was a bit confusing so here I manually entered the years and put them into breaks using cut() function. Initially I thought of using couple of  Year columns  i.e Year1 and Year 2 but that was something I was not convinced of, so eventually I decided to implement this.

Graph
```{r fig.height = 10, fig.width = 10}
graph1<-ggplot(data=cap_growth_decay, aes(x=Year, y=growth, group = Indigenous.region, colour = Indigenous.region)) + geom_line()+ggtitle("Growth of the total Indigenous populations for the capitals of the 8 state/territories across all time periods.")+ ylab("Growth Rate")
graph1

```
As per the plot above the growth rate of Darwin has been gradually declined over the years whereas the max growth rate is in ACT with many ups and downs. However the growth of Melbourne has increased drastically from 1.60 to 2.2 surpassing Perth and Adelaide with a fine margin. Perth and Adelaide does not increase much over time. The interesting fact to observe is that there has been a considerable surge in growth rate of Sydney  from 1.70 to 2.10 from 2016-2031 whereas the growth rate of Tasmania is 1.78 to 2.05
To recapitulate, All the seven states except Darwin witness considerable growth rate of total Indigenous populations for the capitals of the 8 state/territories across all time periods.
```{r}
data_clean1
```









```{r}
head(main)
```


--------------------------------------*****************************A2--------------------------------------------------------------

A2. Investigating the Ages of Indigenous Australians
---------------------------------------------------------------------------------------------------------------------------------

Using Data1, which region has the highest percentage of children in its total 2016
population?
For this, calculate this as a percentage of the total population for a region. The ABS
commonly considers children to be under 15 years of age.

---------------------------------------------------------------------------------------------------------------------------------



To solve this problem I will be performing following things
1)Take the Regionwise data
2)Apply filter on Age column i.e age should be betwen 0-15
3)Apply year filter i.e 2016
4)group the data by region


```{r}

```




```{r}
children_2016<-main%>%
    group_by(Indigenous.region) %>%
    filter(Year == '2016')%>%
    filter(Age == '0 - 4' | Age=='5 - 9' | Age=='10 - 14' )%>%
    summarise(Children = sum(Value, na.rm = TRUE)) %>%
    arrange(Indigenous.region)
children_2016
```

5) Calculate the Total population in the year 2016 for all age groups


```{r}
total_pop_age<-main%>%
    group_by(Indigenous.region) %>%
    filter(Year == '2016')%>%
    summarise(Total_Population = sum(Value, na.rm = TRUE)) %>%
    arrange(Indigenous.region)
total_pop_age
```
6) Merge both the above datasets


```{r}
Total_children<-merge(children_2016,total_pop_age,Key='Indigenous.region')
```
7) Calculate the percentage by dividing number of children with total population in 2016

```{r}
Total_children$percent=Total_children$Children/Total_children$Total_Population*100
Total_children

```
8)using which,max function to extract maximum value
```{r}
Highest_percentage_Children<-Total_children[which.max(Total_children$percent),]
Highest_percentage_Children
```
Therefore the Toowoomba - Roma region has the highest percentage of children in its total 2016 population with 39.76%.



















--------------------------------------------------------------------------------------------------------------------------------

2)
Data2 includes estimated populations measured for the years 2006-2016 and projectedestimates predicted for the years 2016-2031. Data1 just uses projected estimates. UsingData2 only, calculate and discuss which state or territory has the highest percentage of children in its total 2006, 2016 and 2031 populations.

---------------------------------------------------------------------------------------------------------------------------------


To solve this problem I will be performing following steps:
1)Read the dataset
```{r}
data2<-read.csv("FIT5145_S2_2020_A1_Data2_IndigAusPopData_bystate.csv",fileEncoding="UTF-8-BOM")
head(data2)

```
2)Pivoting and performing transformation on dataset to make it understandable

```{r}
data2_df <- pivot_longer(data2, cols=3:28, names_to = "Year", values_to = "Values")
head(data2_df)

```

3)The data after converting to pivot has following problems:
Year has a extra character like X2006. To convert it into correct format I used gsub()function
```{r}
data2_df$Year=gsub("\\X","",data2_df$Year)
```


Remove comma from Values: I used gsub for this scenario as well

```{r}
data2_df$Values=gsub(",","",data2_df$Values)
```
The Age column is in factor format and has values like 65 & above:

Firstly, Convert the age column to character then 
replace the values as 65 and 
finally convert it into numeric format


```{r}
data2_df$Age=as.character(data2_df$Age)
data2_df$Age[data2_df$Age == "65 and over" ] <- "65"
data2_df$Age=as.numeric(data2_df$Age)
```

Convert Values column to numeric format

```{r}
data2_df$Values=as.numeric(data2_df$Values)
head(data2_df)
main2<-data2_df
```



Children 2006



4)creating a new dataframe which stores children data for the year 2006 and age<=15

```{r}
Children_2006<-data2_df%>%
    group_by(State,Year) %>%
    filter(Age <= 15)%>%
    filter(Year=='2006')%>%
    summarise(Children = sum(Values, na.rm = TRUE)) %>%
    arrange(State)
Children_2006

```
5)Calculate total population of the year 2006

```{r}
Total_Population_2006<-data2_df%>%
    group_by(State,Year) %>%
    filter(Year=='2006')%>%
    summarise(Total_Population = sum(Values, na.rm = TRUE)) %>%
    arrange(State)
Total_Population_2006

```
6)merge the above two datasets using State column

```{r}
Total_children_2006<-merge(Children_2006,Total_Population_2006,Key='State')
Total_children_2006
```

7)calculate percentage of children
```{r}
Total_children_2006$percent=Total_children_2006$Children/Total_children_2006$Total_Population*100
Total_children_2006
```
8) calculate the maximum percentage
```{r}
Highest_percentage_Children_2006<-Total_children_2006[which.max(Total_children_2006$percent),]
Highest_percentage_Children_2006 
```
  





children 2016







```{r}
Children_2016<-data2_df%>%
    group_by(State,Year) %>%
    filter(Age <= 15)%>%
    filter(Year=='2016')%>%
    summarise(Children = sum(Values, na.rm = TRUE)) %>%
    arrange(State)
Children_2016

```
Total Population 2016


```{r}
Total_Population_2016<-data2_df%>%
    group_by(State,Year) %>%
    filter(Year=='2016')%>%
    summarise(Total_Population = sum(Values, na.rm = TRUE)) %>%
    arrange(State)
Total_Population_2016

```

merge the above two datasets using State column

```{r}
Total_children_2016<-merge(Children_2016,Total_Population_2016,Key='State')
Total_children_2016
```
Percentages of Children in 2016

```{r}
Total_children_2016$percent=Total_children_2016$Children/Total_children_2016$Total_Population*100
Total_children_2016
```


 calculate the maximum percentage of children for 2016
```{r}
Highest_percentage_Children_2016<-Total_children_2016[which.max(Total_children_2016$percent),]
Highest_percentage_Children_2016 
```







Children for year 2031







```{r}
Children_2031<-data2_df%>%
    group_by(State,Year) %>%
    filter(Age <= 15)%>%
    filter(Year=='2031')%>%
    summarise(Children = sum(Values, na.rm = TRUE)) %>%
    arrange(State)
Children_2031

```

Total Population for 2031

```{r}
Total_Population_2031<-data2_df%>%
    group_by(State,Year) %>%
    filter(Year=='2031')%>%
    summarise(Total_Population = sum(Values, na.rm = TRUE)) %>%
    arrange(State)
Total_Population_2031

```
merge both datasets by State
```{r}
Total_children_2031<-merge(Children_2031,Total_Population_2031,Key='State')
Total_children_2031
```
Children Percentages for 2031
```{r}
Total_children_2031$percent=Total_children_2031$Children/Total_children_2031$Total_Population*100
Total_children_2031
```

 calculate the maximum percentage of children for 2016
```{r}
Highest_percentage_Children_2031<-Total_children_2031[which.max(Total_children_2031$percent),]
Highest_percentage_Children_2031
```












The Highest percentage of children in 2006, 2016 and 2031 is 42.08% 38.16%,33.59% respectively in QLD(Queensland)













---------------------------------------------------------------------------------------------------------------------------------

3. Use R to build a Motion Chart comparing the total Indigenous Australian population of each region to the percentage of Indigenous Australian children in each state/territory. Use the region populations calculated from Data1 and the child percentage values calculated from Data2. The motion chart should show the population on the x-axis, the percentage on the y-axis, the bubble size should depend on the population.
---------------------------------------------------------------------------------------------------------------------------------
```{r}
head(data_clean1)






```

TO solve this question I have used if else loop to add State column in Data1 using the Region Code column 
Secondly I have extracted State,Indigenous_region, Year and Region Code from Data1
Thirdly, I extractd the Children percentage from data2
Finally I have merged both the tables and plotted on motion chart using googleViz library.




```{r}
data1_bubble<-data_clean1
#creates a new column with foolowing values
data1_bubble$State <- ifelse(data1$INDIGENOUS_REGION >= 100 & data1$INDIGENOUS_REGION <= 109, 'NSW', 
                           ifelse(data1$INDIGENOUS_REGION >= 200 & data1$INDIGENOUS_REGION <= 299, 'VIC',
                                  ifelse(data1$INDIGENOUS_REGION >= 300 & data1$INDIGENOUS_REGION <= 399, 'QLD',
                                         ifelse(data1$INDIGENOUS_REGION >= 400 & data1$INDIGENOUS_REGION <= 499, 'SA',
                                                ifelse(data1$INDIGENOUS_REGION >= 500 & data1$INDIGENOUS_REGION <= 599, 'WA',
                                                       ifelse(data1$INDIGENOUS_REGION >= 600 & data1$INDIGENOUS_REGION <= 699, 'TAS',
                                                              ifelse(data1$INDIGENOUS_REGION >= 700 & data1$INDIGENOUS_REGION <= 799, 'NT', 'ACT')))))))

```

```{r}
data1_bubble
```
```{r}
a10<-data1_bubble%>%
  group_by(Region_Code,Indigenous.region,Year,State)%>%
  summarise(TotalPopulation = sum(Value, na.rm = TRUE))
a10
```



region populations calculated from Data1 






Children population grouped by State

```{r}
children_bubble<-main2%>%
    group_by(State,Year) %>%
    filter(Age <= 15)%>%
    filter(Year !='2006'&& Year !='2007' && Year !='2008'&& Year !='2009'&& Year !='2010'&& Year !='2011'&& Year !='2012'&& Year !='2013'&& Year !='2014'&& Year !='2015')%>%
    summarise(Children = sum(Values, na.rm = TRUE)) %>%
    arrange(State)
children_bubble

```
 Total population from data2
```{r}
Tot_Population_bubble<-main2%>%
    group_by(State,Year) %>%
    filter(Year !='2006'&& Year !='2007' && Year !='2008'&& Year !='2009'&& Year !='2010'&& Year !='2011'&& Year !='2012'&& Year !='2013'&& Year !='2014'&& Year !='2015')%>%
    summarise(Total_pop = sum(Values, na.rm = TRUE)) %>%
    arrange(State)
Tot_Population_bubble
```
merge children_bubble and Total_Population_bubble omn  Year

```{r}
Bubble_merged<-merge(children_bubble,Tot_Population_bubble,Key='State')
Bubble_merged
```
Calculate the percentage from above tibble

```{r}
Bubble_merged$percent=(children_bubble$Children/Tot_Population_bubble$Total_pop)*100
Bubble_merged
```


Finally merge both the datasets from data1 and data2


```{r}
final_merge<-merge(a10,Bubble_merged,by=c('Year','State'))
final_merge
```
Finally we can plot the motion chart using :

```{r}
library('googleVis')
Motion=gvisMotionChart(final_merge, idvar="Indigenous.region",timevar='Year',sizevar = 'TotalPopulation',xvar ='TotalPopulation',yvar='percent' )

#display the motion chart
plot(Motion)

```

-------------------------------------------------------------------------------------------------------------------------------
4. Using the Motion Chart, answer the following questions, supporting your answers with
relevant R code and/or Motion Charts
---------------------------------------------------------------------------------------------------------------------------------
a. Which region’s population overtakes that of another region in the same state/territory?
In which year/s does this happen?

-----------------------------------------------------------------------------------------------------------------------------------

1) As per my analysis on data the region Apatula of Northern Territory(NT) overtakes Katherine of Northern Territory(NT) in the year 2021. 


-------------------------------------------------------------------------------------------------------------------------------


b. Is there generally a relationship between the Indigenous Australian population size and
percentage of children in the population? If so, what kind of relationship? Explain
your answer.

---------------------------------------------------------------------------------------------------------------------------------
As per my analysis, As the year progresses from 2016 to 2031 there is a increase in total number of population age(0-65) in different states and regions but, the percentage of children in differnt regions comes down as their age increases which we can clearly see in the motion chart.TO recapitulate as the population increase percentage of children decreases.


--------------------------------------------------------------------------------------------------------------------------------
c.Colour is commonly used in data visualisation to help understand data. Which aspect of
this data would you use colour for in your plot and why?

I would prefer using color for the column called State in the data provided, as the number of values in the column State are less and clearly differentiable which is the main objective of visualisation i.e to understand the plot properly and gain insights from it. If we would have taken Region or Age as a color it would be  much of  a chaos and hard to interpret.


---------------------------------------------------------------------------------------------------------------------------------
d. Are there any other interesting things you notice in the data or any changes you would
recommend for the Motion Chart?

The interesting insight which completely amazed me after visualizing the motion chart was that in the year 2016 region Brisbane(QLD:Total Population 84254) was behind NSW Central and North Coast(NSW: Total Population 85169) but as the years progressed we can notice that in the year 2019 Brisbane started overtaking NSW Central and North Coast. By the year 2031 the population of Brisbane(129835) was more than NSW Central and North Coast(123722).

---------------------------------------------------------------------------------------------------------------------------------




















------------------------------------------------------B-------------------------------------------------------------------












--------------------------------------*****************************--------------------------------------------------------------
B: Exploratory Analysis on Australian Immunisation rates

1. How many PHN areas does the data cover?

```{r}
#read the data
AusImmunisationData <- read.csv('FIT5145_S2_2020_A1_Data3_AusImmunisationData.csv',fileEncoding="UTF-8-BOM")
AusImmunisationData
```
The dataset displays yearly data regarding the number of 1, 2 and 5 year-old Australian children fully or partially immunised in various Primary Health Network (PHN) areas.

After analyzing data the column Number.fully.immunised,Number.not.fully.immunised,Number.of.registered.IndigAus..children,Number.IndigAus.fully.immunised is factor type instead of integer type, 

Moreover columns Number.fully.immunised,Number.not.fully.immunised have commas.

To solve this issue I first tried to convert into numeric directly by using as.numeric(column_name). However, on converting the data also changed which was not appropriate so I did following steps to remove commas and make it of type numeric.

```{r}
#converts factor to numeric
AusImmunisationData$Number.fully.immunised=as.numeric(gsub(",","", AusImmunisationData$Number.fully.immunised))
AusImmunisationData$Number.of.registered.children=as.numeric(gsub(",","", AusImmunisationData$Number.of.registered.children))
AusImmunisationData$Number.of.registered.IndigAus..children=as.numeric(gsub(",","",AusImmunisationData$Number.of.registered.IndigAus..children))
AusImmunisationData$Number.IndigAus.fully.immunised=as.numeric(gsub(",","", AusImmunisationData$Number.IndigAus.fully.immunised))
AusImmunisationData
```

AusImmunisationData



For remaining two columns i.e. Number.of.registered.IndigAus..children,Number.IndigAus.fully.immunised
we can convert it to numeric type using normal conversion function like 

---------------------------------------------------------------------------------------------------------------------------------
B1. Values and Variables

How many PHN areas does the data cover?
---------------------------------------------------------------------------------------------------------------------------------
To answer this question, I passed the PHN.area.name column to data set, to get data Area_wise Group_by is implemented. Moreover, the important part is to find the unique rows in the column, I used n_distinct() in summarise function.
```{r}
phone_area_df<-AusImmunisationData['PHN.area.name']%>%
  group_by(PHN.area.name)
area<-phone_area_df%>%
  summarise_all(funs(n_distinct(.)))
#count the number of rows
nrow(area)
```
The data covers 31 PHN areas


---------------------------------------------------------------------------------------------------------------------------------
2 What are the possible values for 'PHN code'?
---------------------------------------------------------------------------------------------------------------------------------
Using the similar approach as above, I created a dataframe phone_code_df in which I passed AusImmunisationData['PHN.code'] column,using group_by to bring the data in phn.code groups.

Then I created another dataframe phone_code_df to summarise all the distinct columns together


```{r}
phone_code_df<-AusImmunisationData['PHN.code']%>%
  group_by(PHN.code)
 code<-phone_code_df%>%
  summarise_all(funs(n_distinct(.)))
code
```






---------------------------------------------------------------------------------------------------------------------------------
3 For each row, calculate the percentage of Australian children that are fully immunised (this is
the immunisation rate). What are the average, maximum and minimum immunisation rates?
Calculate the same for the group that are Indigenous Australian children. Do all of those
values seem statistically reasonable to you?
--------------------------------------------------------------------------------------------------------------------------------






immuni_rate
```{r}
immuni_rate<-AusImmunisationData%>%
    select(State,Age.group,Reporting.Year,Number.of.registered.children,Number.fully.immunised)
immuni_rate$percent=round(immuni_rate$Number.fully.immunised/immuni_rate$Number.of.registered.children*100,2)
  immuni_rate
```
For calculating the mean, Minimum and maximum values max(),min() and mean() functions are used to calculate the mean maximum and minimum 
```{r}
Highest_percentage_Aus_immuni<-immuni_rate[which.max(immuni_rate$percent),]
Highest_percentage_Aus_immuni
```
The maximum immunisation rate of Aus Childre children that are fully immunised is 96.15%
```{r}
Lowest_percentage_Aus_immuni<-immuni_rate[which.min(immuni_rate$percent),]
Lowest_percentage_Aus_immuni
```
The minimum immunisation rate of Aus children that are fully immunised is 86.1%
```{r}
Average_percentage_Aus_immuni<-mean(immuni_rate$percent)
Average_percentage_Aus_immuni

```
The mean of Average_percentage_Aus_immuni  is 92.06159%










Indigenous Children

To answer this question I had to remove the rows containing the 
```{r}
Immu_rate_Indi<-AusImmunisationData%>%
    select(State,Age.group,Reporting.Year,Number.of.registered.IndigAus..children,Number.IndigAus.fully.immunised)%>%
    filter(Number.of.registered.IndigAus..children!='NP' & Number.IndigAus.fully.immunised != 'NP')
    Immu_rate_Indi$percentage=Immu_rate_Indi$Number.IndigAus.fully.immunised/Immu_rate_Indi$Number.of.registered.IndigAus..children*100

```


```{r}
Highest_percentage_Indi_immuni<-Immu_rate_Indi[which.max(Immu_rate_Indi$percentage),]
Highest_percentage_Indi_immuni

```

The Maximum immunisation rate of Indigenous children that are fully immunised  is 97.68%
```{r}
Lowest_percentage_Indi_immuni<-Immu_rate_Indi[which.min(Immu_rate_Indi$percentage),]
Lowest_percentage_Indi_immuni
```
The minimum immunisation rate of Indigenous children that are fully immunised  is 73.77%
```{r}
Average_percentage_Ind_immuni<-mean(Immu_rate_Indi$percentage)
Average_percentage_Ind_immuni

```
The average immunisation rate of Indigenous children that are fully immunised  is 89.92%




Statistically,the values sounds reasonable as the immunisation rate of Australian Children is more as compared to that of Indigenous Children. The average immunisation rate of 96% clearly states that Australian parents are taking immunising children seriously, On the other hand the average percentage of Indigenous Children is 89.92% which is not bad.The maximum immunisation rate for Australian Children is 96% whereas the interesting factor to observe is that the maximum immunisation rate for Indigenous Children is 97% in NSW.However the minimum immunisation rate of Indigenous children is mere 73%  in WA.







---------------------------------------------------------------------------------------------------------------------------------B2. Variation in rates over Time, Age and Location
Generate boxplots (or other plots) of the immunisation rates versus year and age to
answer the following questions:
---------------------------------------------------------------------------------------------------------------------------------

To construct the boxplots I used (ggplot2) library along with above created dataframes i.e Immu_rate_Indi and immuni_rate dor Indigenous Children and Australian Children respectively
```{r}
library(ggplot2)
main<-Immu_rate_Indi%>%
    select(Age.group,Reporting.Year,percentage)
    
    ggplot(main, aes(x=Reporting.Year, y=percentage, fill=Age.group)) +
    geom_boxplot()+ggtitle("Box plot for Indigenous Children for different age groups")

```
1. Have the immunisation rates improved over time? Are the median immunisation rates
increasing, decreasing or staying the same?
---------------------------------------------------------------------------------------------------------------------------------
The immunisation rate for Indigenous Australian Children has overall has remain constant somewhere around 90%.

--------------------------------------------------------------------------------------------------------------------------------
2. How do the immunisation rates vary with the age of the child?
--------------------------------------------------------------------------------------------------------------------------------
Age wise, for children having age 1 there has been constant increase in immunisation rate i.e from 85% to 92% in 2016-17. Whereas for age group 2 Indigenous Australian Children immunisation hasdeclined drastically from 91% to 88% in 2016-17.However the immunisation rate for age group 5 was constant from 2012-2014 and after that it spiked to 95% in 2016-2017.


```{r}
age_rates_Ind<-Immu_rate_Indi%>%
     group_by(Reporting.Year,Age.group) %>%
     summarise(Median=median(percentage))%>%
     arrange((Age.group))
age_rates_Ind

```
Australian Children boxplot
```{r}
library(ggplot2)
main1<-immuni_rate%>%
    select(Age.group,Reporting.Year,percent)
    
    ggplot(main1, aes(y=percent,x=Reporting.Year, fill=Age.group)) +
    geom_boxplot()+ggtitle("Box plot for Australian Children for different age groups")

```
Answer 1 for Australian Children
---------------------------------------------------------------------------------------------------------------------------------
The immunisation rate for  Australian Children has overall has descreased gradually every year i.e from 91% to 93%.
--------------------------------------------------------------------------------------------------------------------------------

Answer 2 Australian Children
--------------------------------------------------------------------------------------------------------------------------------
Age.group wise, for children having age 1 there has been constant increase in immunisation rate i.e from 91% to 94% in 2016-17. Whereas for age group 2  Australian Children immunisation is fluctuating i.e it increased from 2012-2014 but after that it decreased and reached 89% in 2014-2015.However the immunisation rate for age group 5 was constant from 2012-2015 and after that it started to increase and reached 94% in 2016-2017.
```{r}
age_rates_Aus<-immuni_rate%>%
     group_by(Reporting.Year,Age.group) %>%
     summarise(Median=median(percent))%>%
     arrange(Age.group)
age_rates_Aus
```

Generate boxplots (or other plots) of the immunisation rates versus locations and answer the following questions:

For generating these boxplots I used ggplot2 library along with geom_boxplot()function. I have used the Immu_rate_Indi dataframe from above to plot this.Moreover,as per the question we need immunisation rates versus locations so I am grouping it up with State.
-------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------
Indigenous Australian Children

```{r}


state_rates<-Immu_rate_Indi%>%
        select(State,percentage)%>%
        group_by(State)
        ggplot(state_rates, aes(x=percentage, y=State, fill=State)) +
    geom_boxplot()+ggtitle("Indigenous Australian Children immunisation rate State wise")

```
3 What is the median rate per state/territory?
---------------------------------------------------------------------------------------------------------------------------------
The median immunisation rate per states is around 90+ for ACT,NSW,NT,QLD,TAS and VIC whereas SA, VIC/NSW,WA has immunisation rate less than 90.The NT being the highest with 93% and WA being the lowest.
---------------------------------------------------------------------------------------------------------------------------------
4.
--------------------------------------------------------------------------------------------------------------------------------
As per my observations, TAS(Tasmania) is more consistent as the range starts from above 90% and it goes upto 93% which clearly states that the immunisation rate is always consistently above 90%.

-------------------------------------------------------------------------------------------------------------------------------
--------------------------Indigenous Australian State boxplot----------------------------------------------------------
--------------------------------------------------------------------------------------------------------------------------------

```{r}
state_rates<-immuni_rate%>%
        select(State,percent)%>%
        group_by(State)
        ggplot(state_rates, aes(x=percent, y=State, fill=State)) +
    geom_boxplot()+ggtitle("Australian Children immunisation rate State wise")



```
Answer to q3 for Australian Children immunisation rate State wise
--->

ACT and VIC/NSW has the same immunisation rate of arounf 93% whereas the WA being the lowest with around 90%.However NT and Sa are in the range of 91% and VIC,QLD,TAS,NSW are in range of 92% 

---------------------------------------------------------------------------------------------------------------------------------
Answer to q4
---------------------------------------------------------------------------------------------------------------------------------

Here ACT is more consistent as it starts at higher range and it flickers around that. Moreover there have not been extreme fall or rise in the immunisation rate for ACT.The range starts at around 92% and moves around that.
